\textbf{The Illusion Game: A Novel Experimental Paradigm Provides Evidence for a General Factor of Visual Illusion Sensitivity and Personality Correlates}

### 2023-01-01 submission/R1/manuscript.Rmd
more

The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and response time (RT) **separately for each illusion**. We started by fitting General Additive Models (GAMs), which can parsimoniously accommodate possible non-linear effects and interactions. Errors were analyzed using Bayesian logistic mixed models, and RTs of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location $\mu$ (mean), scale $\sigma$ (spread) and tail-dominance $\tau$ of the RT distribution [@balota2011moving; @matzke2009psychological].

Using GAMs as the "ground-truth" models, we attempted at approximating them using general linear mixed models, which can be used to estimate the effects' participant-level variability (via random slopes). Following a comparison of models with a combination of transformations (raw, log, square root or cubic root) on the main predictors (task *difficulty* and illusion *strength*), we fitted the best model (based on their indices of fit), and compared their output visually (**Figure 3**).

The inter-individual variability in the effect of illusion strength and its interaction with task difficulty (**diff**) was extracted from the models and used as participant-level scores. We then explored the relationship of these indices across different illusions using exploratory factor analysis (EFA) and structural equation modelling (SEM), and tested the existence of a general factor of illusion sensitivity (Factor *i*).

>>>>

The first part of the analysis focused on modelling the effect of illusion strength and task difficulty on errors and response time (RT) **separately for each illusion** under a Bayesian framework. We started by fitting General Additive Models (GAMs), which can parsimoniously accommodate possible non-linear effects and interactions. Errors were analyzed using logistic mixed models **(suited to estimate the error rate)**, and RTs of correct responses were analyzed using an ex-Gaussian family with the same fixed effects entered for the location $\mu$ (mean), scale $\sigma$ (spread) and tail-dominance $\tau$ of the RT distribution [@balota2011moving; @matzke2009psychological].

Using GAMs as the "ground-truth" models, we attempted at approximating them using general linear mixed models, which can be used to estimate the effects' participant-level variability (via random slopes). Following a comparison of models with a combination of transformations (raw, log, square root or cubic root; **which are types of relationship commonly found in perceptual tasks**) on the main predictors (task *difficulty* and illusion *strength*), we fitted the best model (**based on their BIC and R2**), and compared their output visually (**Figure 3**). **Note that the model comparison and the parameters used in the resulting models were not pre-registered.**

The inter-individual variability in the effect of illusion strength and its interaction with task difficulty (**diff**) was extracted from the models and used as participant-level scores. We then explored the relationship of these indices across different illusions using exploratory factor analysis (EFA, **to gain insights into the structure**), and structural equation modelling (SEM, **to model and test different hierarchical models**), and tested the existence of a general factor of illusion sensitivity (Factor *i*).
